Full Dissertation:

### 2024-04-04 Sections/full.tex
fixes

---
title: Full Dissertation
subtitle: An instance-based model account of the benefits of varied practice in visuomotor skill
date: 04/04/24
#page-layout: full
code-fold: true
code-tools: true
toc: true
toc-location: body
toc-title: Contents
execute: 
  warning: false
  eval: true
format:
  html: default
  hugo-md:
    echo: false
    output: asis
    html-math-method: mathjax
    output-file: diss_full_hugo.md
  gfm:
    echo: true
    output-file: diss_full_gfm.md
  hikmah-manuscript-pdf:
    echo: false
    output-file: "hikmah-e1.pdf"
    mainfont: "Linux Libertine O"
    mainfontoptions:
      - "Numbers=Proportional"
      - "Numbers=OldStyle"
    mathfont: "Libertinus Math"
editor: 
  markdown: 
    wrap: 72
prefer-html: true
---

# Variability and Generalization

@@

## The study of variability

Studies investigating the "benefits of variability" hypothesis usually assign participants to either a constant or varied group for the training stage of the experiment. Then, subjects in both groups complete an identical testing stage which often consists items/conditions seen during training, and novel items/conditions. If the varied group performs better in the testing stage, this is taken for evidence of the benefits of variability hypothesis. Even within this relatively straightforward between-groups design, researchers must navigate several crucial methodological choices, highlighted below:

1)  Variables Subject to Variation. In multidimensional tasks, researchers have the option to vary numerous variables. The experimenters must decide the specific dimension(s) across which variation will occur. For instance, in a projectile throwing accuracy task -- researchers might vary the distance from the target, the size of the target, the weight of the projectile. They might also vary a contextual variable not directly relevant to the task, but which will still be encoded by the subject on a trial by trial basis, e.g. the background color.

2)  Magnitude of Variation, relative to the control condition. The simplest comparison would be to compare a constant group who trains with 1 example/condition, against a varied group that trains from 2 examples/conditions. However, it is not uncommon in the literature for the varied condition to train from 3 or 4 conditions. For example, @catalanoDistantTransferCoincident1984a train varied subjects from 4 different velocities in their coincident timing task, and [@goodeSuperiorityVariableRepeated2008] have varied subjects' practice with 3 different variants (i.e. different letter scrambles of the same word) of an anagram for a given word, while their constant participants view the same variant 3 different times. Alternatively, rather than a constant vs. varied comparison, subjects in all conditions might experience a variety of training items, but with one group experiencing a greater number of unique items [@nosofskyModelguidedSearchOptimal2018].

3)  Locations within Task-Space. For tasks in which the stimuli or conditions fall within a continuous metric space, the experimenter must decide whether the varied instances are relatively close together (e.g. throwing a ball from a distance of 4 feet and 5 feet), or far apart (throwing from 4 feet and 20 feet). Spreading the varied training items further apart may be beneficial in terms of providing a more representative sample of the task space to the learner, however large distances may also result in significant differences in difficulty between the training examples, which can be a common confound in variability studies.

4)  Proximity of Testing to Training Conditions. Intuitively, the fairest form of comparison is to include testing conditions that are of an equivalent distance from both the varied and constant groups. However researchers might also attempt to demonstrate the benefits of variation as being sufficiently powerful to outperform constant training, even in cases where the constant group trained from a closer proximity to the testing conditions, or whose training conditions are identical to the testing conditions [@goodeSuperiorityVariableRepeated2008; @kerrSpecificVariedPractice1978].

## Variability Literature Review

An early and influential work on the influence of variability on category learning is that of @posnerGenesisAbstractIdeas1968. In an ambitious attempt to address the question of how category information is represented, the authors trained participants to categorize artificial dot patterns, manipulating whether learners were exposed to examples clustered close to the category prototypes (e.g. a low variability condition), or spread further away from the prototype (the varied-training group). It should be noted that both groups in this study were trained with the same number of unique instances and the manipulated difference was how spread out the instances were. The authors claim based on prior experiments using the same stimuli, that the training stimuli for the varied group were at least as far away from the testing stimuli as the training stimuli of the less-varied group. The authors interpreted their findings as evidence for the extraction of an abstraction or schema that is extracted and stored, and then over time becomes more likely to be the reference point from which generalization occurs, given that specific instances are thought to decay at a faster rate than prototypes or schema. The Posner and Keele study has been extremely influential and continues to be cited in contemporary research as clear evidence that schema abstraction underlies the benefits of varied training. It's also referenced as a key influence in the development of "Schema Theory of Motor Learning" [@schmidtSchemaTheoryDiscrete1975], which in turn influenced decades of investigations on the potential benefits of varied training in motor skill learning. However, the classic Posner & Keele study despite being far more carefully designed than many subsequent studies, and despite being a relative rarity in explicitly discussing and attempting to control for potential confounds of similarity between groups, may nevertheless be emblematic of a common issue in many investigations of the effects of varied training on learning. The problem with Posner & Keele's conclusion was demonstrated clearly almost 3 decades later [@palmeriCentralTendenciesExtreme2001], when researchers conducting a near replication of the original study also collected similarity judgements following training and performed multidimensional scaling analysis. Rather than being in the middle of the training stimuli as was the case in the physical stimuli space, the psychological representation of the prototype was shown reside at an extreme point, and generalization patterns by participants that would have seemed to warrant the learning of a prototype were then easily accounted for with only the assumption that the participants encoded instances. One of the primary concerns of the present paper is that many of the studies which purport to explain the benefits of variation via prototypes, schemas, or other abstractions, are often overlooking the potential of instance based similarity accounts.

## Motor skill learning

Training variation has also been shown to promote transfer in motor learning. Much of this research has been influenced by the work of [@schmidtSchemaTheoryDiscrete1975], who proposed a schema-based account of motor learning as an attempt to address the longstanding problem of how novel movements are produced. Schema theory presumes a priori that learners possess general motor programs for classes of movements, such as an underhand throw. When called up for use, such programs must be parameterized, as well as schema rules that determine how a motor program is parameterized or scaled for a particular movement. Schema theory predicts that varied training results in the formation of a more general schema-rule, which can allow for transfer to novel movements within a given movement class, such as an underhand throw (though it is agnostic to the development of the movement classes themselves). Experiments that test this hypothesis are often designed to compare the transfer performance of a constant-trained group against that of a varied-trained group. Both groups train on the same task, but the varied group practices with multiple instances along some task-relevant dimension that remains invariant for the constant group. For example, investigators might train two groups of participants to throw a projectile at a target, with a constant group that throws from a single location, and a varied group that throws from multiple locations. Both groups are then tested from novel locations.

One of the earliest, and still often cited investigations of Schmidt's benefits of variability hypothesis was the work of @kerrSpecificVariedPractice1978. Two groups of children, aged 8 and 12, were assigned to either constant or varied training of a bean bag throwing task. The constant group practiced throwing a bean-bag at a small target placed 3 feet in front of them, and the varied group practiced throwing from a distance of both 2 feet and 4 feet. Participants were blindfolded and unable to see the target while making each throw but would receive feedback by looking at where the beanbag had landed in between each training trial. 12 weeks later, all of the children were given a final test from a distance of 3 feet which was novel for the varied participants and repeated for the constant participants. Participants were also blindfolded for testing and did not receive trial by trial feedback in this stage. However, at the halfway point of the testing stage they were allowed to see the landing location of the 4 beanbags they had thrown, and then completed the final 4 testing throws. In both age groups, participants performed significantly better in the varied condition than the constant condition, though the effect was larger for the younger, 8-year-old children. Although this design does not directly assess the hypothesis of varied training producing superior generalization to constant training (since the constant group is not tested from a novel position), it nevertheless offers a compelling example of the merits of varied practice.

On occasion the Kerr and Booth design may be nested within a larger experimental design. One such study that used a movement timing task, wherein subjects had to move their hand from a starting location, to a target location, attempting to arrive at the target location at specific time following the onset of a cue [@wrisbergVariabilityPracticeHypothesis1987]. This study utilized 4 different constant groups, and 3 varied groups, with one of the constant groups training under conditions identical to the testing conditions, and which were not trained on by any of the varied groups, e.g. the design of Kerr and Booth. However, in this case the varied group did not outperform the constant group. A more recent study attempting a slightly more direct replication of the original Kerr & Booth study [@willeyLongtermMotorLearning2018], having subjects throw beanbags at a target, with the varied group training from positions (5 and 9 feet) on either side of the constant group (7 feet). However, this study diverged from the original in that the participants were adults; they faced away from the target and threw the beanbag backwards over their bodies; they alternated using their right and left hands every 6 trials; and underwent a relatively extreme amount of training (20 sessions with 60 practice trials each, spread out over 5-7 weeks). Like @wrisbergVariabilityPracticeHypothesis1987, this study did not find a varied advantage from the constant training position, though the varied group did perform better at distances novel to both groups.

Some support for the Kerr and Booth findings was found with a relatively less common experimental task of training participants in hitting a projectile at a target with the use of a racket [@greenPracticeVariabilityTransfer1995a]. Varied participants trained with tennis, squash, badminton, and short-tennis rackets were compared against constant subjects trained with only a tennis racket. One of the testing conditions had subjects repeat the use of the tennis racket, which had been used on all 128 training trials for the constant group, and only 32 training trials for the varied group. Nevertheless, the varied group outperformed the constant group when using the tennis racket at testing, and also performed better in conditions with several novel racket lengths. Of course, this finding is less surprising than that of Kerr & Booth, given that varied subjects did have some prior exposure to the constant groups condition. This highlights an issue rarely discussed in the literature, of how much practice from an additional position might be necessary to induce benefits. Experimenters almost uniformly have varied participants train with an equivalent number of trials from each of their conditions

One of the few studies that has replicated the surprising result of varied outperforming constant, from the constant training condition, did so in the relatively distant domain of verbal manipulation [@goodeSuperiorityVariableRepeated2008]. All participants trained to solve anagrams of 40 different words ranging in length from 5 to 11 letters, with an anagram of each word repeated 3 times throughout training, for a total of 120 training trials. Although subjects in all conditions were exposed to the same 40 unique words (i.e. the solution to an anagram), participants in the varied group saw 3 different arrangements for each solution-word, such as DOLOF, FOLOD, and OOFLD for the solution word FLOOD, whereas constant subjects would train on three repetitions of LDOOF (spread evenly across training). Two different constant groups were used. Both constant groups trained with three repetitions of the same word scramble, but for constant group A, the testing phase consisted of the identical letter arrangement to that seen during training (e.g. LDOOF), whereas for constant group B, the testing phase consisted of a arrangement they had not seen during training, thus presenting them with a testing situation similar situation to the varied group. At the testing stage, the varied group outperformed both constant groups, a particularly impressive result, given that constant group A had 3 prior exposures to the word arrangement (i.e. the particular permutation of letters) which the varied group had not explicitly seen. However varied subjects in this study did not exhibit the typical decrement in the training phase typical of other varied manipulations in the literature, and actually achieved higher levels of anagram solving accuracy by the end of training than either of the constant groups -- solving 2 more anagrams on average than the constant group. This might suggest that for tasks of this nature where the learner can simply get stuck with a particular word scramble, repeated exposure to the identical scramble might be less helpful towards finding the solution than being given a different arrangement of the same letters. This contention is supported by the fact that constant group A, who was tested on the identical arrangement as they experienced during training, performed no better at testing than did constant group B, who had trained on a different arrangement of the same word solution -- further suggesting that there may not have been a strong identity advantage in this task.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  letterpaper,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
  \setmainfont[Numbers=Proportional,Numbers=OldStyle]{Linux Libertine O}
  \setmathfont[]{Libertinus Math}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{xcolor}
\usepackage[top=1in,bottom=1in,left=1in,right=1in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% -----------------------
% CUSTOM PREAMBLE STUFF
% -----------------------

% -----------------
% Typography tweaks
% -----------------
% Indent size
\setlength{\parindent}{0.5in}
\setlength{\leftmargin}{0.5in}

% Fix widows and orphans
\usepackage[all,defaultlines=2]{nowidow}

% List things
\usepackage{enumitem}
% Same document-level indentation for ordered and ordered lists
\setlist[1]{labelindent=\parindent}
\setlist[itemize]{leftmargin=*}
\setlist[enumerate]{leftmargin=*}

% Wrap definition list terms
% https://tex.stackexchange.com/a/9763/11851
\setlist[description]{style=unboxed}

% For better TOCs
\usepackage{tocloft}

% Remove left margin in lists inside longtables
% https://tex.stackexchange.com/a/378190/11851
\AtBeginEnvironment{longtable}{\setlist[itemize]{nosep, wide=0pt, leftmargin=*, before=\vspace*{-\baselineskip}, after=\vspace*{-\baselineskip}}}

% For fancy ORCID links
\usepackage{orcidlink}

% Indent all first paragraphs because APA wants that
\usepackage{indentfirst}

% -----------------
% Title block stuff
% -----------------

% Abstract
\usepackage{abstract}
\renewcommand{\abstractnamefont}{\normalfont\normalsize\bfseries}
\renewcommand{\abstracttextfont}{\normalsize}
\setlength{\absleftindent}{\parindent}
\setlength{\absrightindent}{\parindent}

% Keywords
\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}}#1}

% Title
\usepackage{titling}
\setlength{\droptitle}{2\baselineskip}
\pretitle{\begin{center}\normalfont\normalsize\bfseries}
\posttitle{\par\end{center}\vspace{2\baselineskip}}

% ------------------
% Section headings
% ------------------
\usepackage{titlesec}
\titleformat*{\section}{\center\normalsize\bfseries}
\titleformat*{\subsection}{\normalsize\bfseries}
\titleformat*{\subsubsection}{\normalsize\bfseries\itshape}
\titleformat{\paragraph}[runin]{\bfseries}{\theparagraph.}{3pt}{\space}[.]
\titleformat{\subparagraph}[runin]{\bfseries\itshape}{\thesubparagraph.}{3pt}{\space}[.]

% \titlespacing{<command>}{<left>}{<before-sep>}{<after-sep>}
% Starred version removes indentation in following paragraph
\titlespacing*{\section}{0em}{2em}{0em}
\titlespacing*{\subsection}{0em}{1em}{0em}
\titlespacing*{\subsubsection}{0em}{0em}{0em}
\titlespacing*{\paragraph}{\parindent}{1ex}{1em}
\titlespacing*{\subparagraph}{\parindent}{1ex}{1em}

% -----------
% Footnotes
% -----------
% NB: footmisc has to come after biblatex because of conflicts
\usepackage[bottom]{footmisc}
\renewcommand*{\footnotelayout}{\footnotesize}

\addtolength{\skip\footins}{12pt}  % vertical space between rule and main text
\setlength{\footnotesep}{16pt}  % vertical space between footnotes

% ----------
% Captions
% ----------
\usepackage[font=normalsize]{caption}

% ---------------------------
% END CUSTOM PREAMBLE STUFF
% ---------------------------
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{caption}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{sidenotes}{}{\usepackage{sidenotes}}
\@ifpackageloaded{marginnote}{}{\usepackage{marginnote}}
\makeatother
\makeatletter
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Full Dissertation},
  colorlinks=true,
  linkcolor={DarkSlateBlue},
  filecolor={Maroon},
  citecolor={DarkSlateBlue},
  urlcolor={DarkSlateBlue},
  pdfcreator={LaTeX via pandoc}}

% -----------------------
% END-OF-PREAMBLE STUFF
% -----------------------

% For patching commands like \subtitle
\usepackage{etoolbox}

% -----------------
% Running headers
% -----------------

\usepackage{fancyhdr}
\setlength{\headheight}{0.25in}
\renewcommand{\headrulewidth}{0pt}  % Remove lines
\renewcommand{\footrulewidth}{0pt}

% SHORT TITLE           Page #
\fancypagestyle{normal}{
  \fancyhf{}
  \lhead{\uppercase{ Full Dissertation }}
  \rhead{\thepage}
}

% Running head: SHORT TITLE        Page #
\fancypagestyle{title}{
  \fancyhf{}
  \lhead{Running head: \uppercase{ Full Dissertation }}
  \rhead{}
}

% Use regular heading style
\pagestyle{normal}

% ----------
% BibLaTeX
% ----------
\usepackage[style=apa,backend=biber]{biblatex}

\setlength\bibitemsep{0pt}  % No space between bib entries
\setlength\bibhang{\parindent}  % Match document indentation

% Fix biblatex's odd preference for using In: by default.
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{%
  \printtext{\bibstring{}\intitlepunct}}}

\addbibresource{../Assets/Bib/Dissertation.bib}

% Start bibliography on new page
\pretocmd{\printbibliography}{\clearpage}{}{}

% ---------------------- 
% Title block elements
% ---------------------- 
\usepackage{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{ and }
\renewcommand*{\Authands}{, and }
\renewcommand*{\Affilfont}{\normalsize}
\renewcommand*{\Authfont}{\normalsize}

\title{Full Dissertation:}

\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par #1 \par}{}{}
}
\makeatother
\subtitle{An instance-based model account of the benefits of varied
practice in visuomotor skill}

\date{}

% Typeset URLs in the same font as their parent environment
%
% This has to come at the end of the preamble, after any biblatex stuff because 
% some biblatex styles (like APA) define their own \urlstyle{}
\usepackage{url}
\urlstyle{same}

% ---------------------------
% END END-OF-PREAMBLE STUFF
% ---------------------------
\begin{document}
% ---------------
% TITLE SECTION
% ---------------

\begin{titlepage}
\center

% Don't include a newpage before \maketitle
{\let\newpage\relax\maketitle}

% Needs to come after \maketitle
\thispagestyle{title}

\vspace{0.25in}

\vspace{0.5in}

\begin{center}

\textbf{Author note}

\end{center}

\vfill  % Fill the rest of the page with whitespace

\end{titlepage}

\doublespacing

\begin{center}
\singlespacing
\textbf{Full Dissertation:\\An instance-based model account of the
benefits of varied practice in visuomotor skill}
\end{center}

% -------------------
% END TITLE SECTION
% -------------------

\renewcommand*\contentsname{Contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\setstretch{2}
\section{Variability and
Generalization}\label{variability-and-generalization}

@@

\subsection{The study of variability}\label{the-study-of-variability}

Studies investigating the ``benefits of variability'' hypothesis usually
assign participants to either a constant or varied group for the
training stage of the experiment. Then, subjects in both groups complete
an identical testing stage which often consists items/conditions seen
during training, and novel items/conditions. If the varied group
performs better in the testing stage, this is taken for evidence of the
benefits of variability hypothesis. Even within this relatively
straightforward between-groups design, researchers must navigate several
crucial methodological choices, highlighted below:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  Variables Subject to Variation. In multidimensional tasks, researchers
  have the option to vary numerous variables. The experimenters must
  decide the specific dimension(s) across which variation will occur.
  For instance, in a projectile throwing accuracy task -- researchers
  might vary the distance from the target, the size of the target, the
  weight of the projectile. They might also vary a contextual variable
  not directly relevant to the task, but which will still be encoded by
  the subject on a trial by trial basis, e.g.~the background color.
\item
  Magnitude of Variation, relative to the control condition. The
  simplest comparison would be to compare a constant group who trains
  with 1 example/condition, against a varied group that trains from 2
  examples/conditions. However, it is not uncommon in the literature for
  the varied condition to train from 3 or 4 conditions. For example,
  \textcite{catalanoDistantTransferCoincident1984a} train varied
  subjects from 4 different velocities in their coincident timing task,
  and \autocite{goodeSuperiorityVariableRepeated2008} have varied
  subjects' practice with 3 different variants (i.e.~different letter
  scrambles of the same word) of an anagram for a given word, while
  their constant participants view the same variant 3 different times.
  Alternatively, rather than a constant vs.~varied comparison, subjects
  in all conditions might experience a variety of training items, but
  with one group experiencing a greater number of unique items
  \autocite{nosofskyModelguidedSearchOptimal2018}.
\item
  Locations within Task-Space. For tasks in which the stimuli or
  conditions fall within a continuous metric space, the experimenter
  must decide whether the varied instances are relatively close together
  (e.g.~throwing a ball from a distance of 4 feet and 5 feet), or far
  apart (throwing from 4 feet and 20 feet). Spreading the varied
  training items further apart may be beneficial in terms of providing a
  more representative sample of the task space to the learner, however
  large distances may also result in significant differences in
  difficulty between the training examples, which can be a common
  confound in variability studies.
\item
  Proximity of Testing to Training Conditions. Intuitively, the fairest
  form of comparison is to include testing conditions that are of an
  equivalent distance from both the varied and constant groups. However
  researchers might also attempt to demonstrate the benefits of
  variation as being sufficiently powerful to outperform constant
  training, even in cases where the constant group trained from a closer
  proximity to the testing conditions, or whose training conditions are
  identical to the testing conditions
  \autocite{goodeSuperiorityVariableRepeated2008,kerrSpecificVariedPractice1978}.
\end{enumerate}

\subsection{Variability Literature
Review}\label{variability-literature-review}

An early and influential work on the influence of variability on
category learning is that of \textcite{posnerGenesisAbstractIdeas1968}.
In an ambitious attempt to address the question of how category
information is represented, the authors trained participants to
categorize artificial dot patterns, manipulating whether learners were
exposed to examples clustered close to the category prototypes (e.g.~a
low variability condition), or spread further away from the prototype
(the varied-training group). It should be noted that both groups in this
study were trained with the same number of unique instances and the
manipulated difference was how spread out the instances were. The
authors claim based on prior experiments using the same stimuli, that
the training stimuli for the varied group were at least as far away from
the testing stimuli as the training stimuli of the less-varied group.
The authors interpreted their findings as evidence for the extraction of
an abstraction or schema that is extracted and stored, and then over
time becomes more likely to be the reference point from which
generalization occurs, given that specific instances are thought to
decay at a faster rate than prototypes or schema. The Posner and Keele
study has been extremely influential and continues to be cited in
contemporary research as clear evidence that schema abstraction
underlies the benefits of varied training. It's also referenced as a key
influence in the development of ``Schema Theory of Motor Learning''
\autocite{schmidtSchemaTheoryDiscrete1975}, which in turn influenced
decades of investigations on the potential benefits of varied training
in motor skill learning. However, the classic Posner \& Keele study
despite being far more carefully designed than many subsequent studies,
and despite being a relative rarity in explicitly discussing and
attempting to control for potential confounds of similarity between
groups, may nevertheless be emblematic of a common issue in many
investigations of the effects of varied training on learning. The
problem with Posner \& Keele's conclusion was demonstrated clearly
almost 3 decades later \autocite{palmeriCentralTendenciesExtreme2001},
when researchers conducting a near replication of the original study
also collected similarity judgements following training and performed
multidimensional scaling analysis. Rather than being in the middle of
the training stimuli as was the case in the physical stimuli space, the
psychological representation of the prototype was shown reside at an
extreme point, and generalization patterns by participants that would
have seemed to warrant the learning of a prototype were then easily
accounted for with only the assumption that the participants encoded
instances. One of the primary concerns of the present paper is that many
of the studies which purport to explain the benefits of variation via
prototypes, schemas, or other abstractions, are often overlooking the
potential of instance based similarity accounts.

\subsection{Motor skill learning}\label{motor-skill-learning}

Training variation has also been shown to promote transfer in motor
learning. Much of this research has been influenced by the work of
\autocite{schmidtSchemaTheoryDiscrete1975}, who proposed a schema-based
account of motor learning as an attempt to address the longstanding
problem of how novel movements are produced. Schema theory presumes a
priori that learners possess general motor programs for classes of
movements, such as an underhand throw. When called up for use, such
programs must be parameterized, as well as schema rules that determine
how a motor program is parameterized or scaled for a particular
movement. Schema theory predicts that varied training results in the
formation of a more general schema-rule, which can allow for transfer to
novel movements within a given movement class, such as an underhand
throw (though it is agnostic to the development of the movement classes
themselves). Experiments that test this hypothesis are often designed to
compare the transfer performance of a constant-trained group against
that of a varied-trained group. Both groups train on the same task, but
the varied group practices with multiple instances along some
task-relevant dimension that remains invariant for the constant group.
For example, investigators might train two groups of participants to
throw a projectile at a target, with a constant group that throws from a
single location, and a varied group that throws from multiple locations.
Both groups are then tested from novel locations.

One of the earliest, and still often cited investigations of Schmidt's
benefits of variability hypothesis was the work of
\textcite{kerrSpecificVariedPractice1978}. Two groups of children, aged
8 and 12, were assigned to either constant or varied training of a bean
bag throwing task. The constant group practiced throwing a bean-bag at a
small target placed 3 feet in front of them, and the varied group
practiced throwing from a distance of both 2 feet and 4 feet.
Participants were blindfolded and unable to see the target while making
each throw but would receive feedback by looking at where the beanbag
had landed in between each training trial. 12 weeks later, all of the
children were given a final test from a distance of 3 feet which was
novel for the varied participants and repeated for the constant
participants. Participants were also blindfolded for testing and did not
receive trial by trial feedback in this stage. However, at the halfway
point of the testing stage they were allowed to see the landing location
of the 4 beanbags they had thrown, and then completed the final 4
testing throws. In both age groups, participants performed significantly
better in the varied condition than the constant condition, though the
effect was larger for the younger, 8-year-old children. Although this
design does not directly assess the hypothesis of varied training
producing superior generalization to constant training (since the
constant group is not tested from a novel position), it nevertheless
offers a compelling example of the merits of varied practice.

On occasion the Kerr and Booth design may be nested within a larger
experimental design. One such study that used a movement timing task,
wherein subjects had to move their hand from a starting location, to a
target location, attempting to arrive at the target location at specific
time following the onset of a cue
\autocite{wrisbergVariabilityPracticeHypothesis1987}. This study
utilized 4 different constant groups, and 3 varied groups, with one of
the constant groups training under conditions identical to the testing
conditions, and which were not trained on by any of the varied groups,
e.g.~the design of Kerr and Booth. However, in this case the varied
group did not outperform the constant group. A more recent study
attempting a slightly more direct replication of the original Kerr \&
Booth study \autocite{willeyLongtermMotorLearning2018}, having subjects
throw beanbags at a target, with the varied group training from
positions (5 and 9 feet) on either side of the constant group (7 feet).
However, this study diverged from the original in that the participants
were adults; they faced away from the target and threw the beanbag
backwards over their bodies; they alternated using their right and left
hands every 6 trials; and underwent a relatively extreme amount of
training (20 sessions with 60 practice trials each, spread out over 5-7
weeks). Like \textcite{wrisbergVariabilityPracticeHypothesis1987}, this
study did not find a varied advantage from the constant training
position, though the varied group did perform better at distances novel
to both groups.

Some support for the Kerr and Booth findings was found with a relatively
less common experimental task of training participants in hitting a
projectile at a target with the use of a racket
\autocite{greenPracticeVariabilityTransfer1995a}. Varied participants
trained with tennis, squash, badminton, and short-tennis rackets were
compared against constant subjects trained with only a tennis racket.
One of the testing conditions had subjects repeat the use of the tennis
racket, which had been used on all 128 training trials for the constant
group, and only 32 training trials for the varied group. Nevertheless,
the varied group outperformed the constant group when using the tennis
racket at testing, and also performed better in conditions with several
novel racket lengths. Of course, this finding is less surprising than
that of Kerr \& Booth, given that varied subjects did have some prior
exposure to the constant groups condition. This highlights an issue
rarely discussed in the literature, of how much practice from an
additional position might be necessary to induce benefits. Experimenters
almost uniformly have varied participants train with an equivalent
number of trials from each of their conditions

One of the few studies that has replicated the surprising result of
varied outperforming constant, from the constant training condition, did
so in the relatively distant domain of verbal manipulation
\autocite{goodeSuperiorityVariableRepeated2008}. All participants
trained to solve anagrams of 40 different words ranging in length from 5
to 11 letters, with an anagram of each word repeated 3 times throughout
training, for a total of 120 training trials. Although subjects in all
conditions were exposed to the same 40 unique words (i.e.~the solution
to an anagram), participants in the varied group saw 3 different
arrangements for each solution-word, such as DOLOF, FOLOD, and OOFLD for
the solution word FLOOD, whereas constant subjects would train on three
repetitions of LDOOF (spread evenly across training). Two different
constant groups were used. Both constant groups trained with three
repetitions of the same word scramble, but for constant group A, the
testing phase consisted of the identical letter arrangement to that seen
during training (e.g.~LDOOF), whereas for constant group B, the testing
phase consisted of a arrangement they had not seen during training, thus
presenting them with a testing situation similar situation to the varied
group. At the testing stage, the varied group outperformed both constant
groups, a particularly impressive result, given that constant group A
had 3 prior exposures to the word arrangement (i.e.~the particular
permutation of letters) which the varied group had not explicitly seen.
However varied subjects in this study did not exhibit the typical
decrement in the training phase typical of other varied manipulations in
the literature, and actually achieved higher levels of anagram solving
accuracy by the end of training than either of the constant groups --
solving 2 more anagrams on average than the constant group. This might
suggest that for tasks of this nature where the learner can simply get
stuck with a particular word scramble, repeated exposure to the
identical scramble might be less helpful towards finding the solution
than being given a different arrangement of the same letters. This
contention is supported by the fact that constant group A, who was
tested on the identical arrangement as they experienced during training,
performed no better at testing than did constant group B, who had
trained on a different arrangement of the same word solution -- further
suggesting that there may not have been a strong identity advantage in
this task.
